---
title: "Adolescent Pregnancies in Kenya (2016 - 2020)"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    source_code: https://github.com/SanmariV/adolescent-pregnancies-kenya
    social: menu
runtime: shiny
resource_files:
- Teenage Pregnancies Data 2016 to 2020_WardLevel_Quarterly.csv
- shared_code.R
---

```{r setup, include=FALSE}
library(flexdashboard)
library(DT)
library(shiny)
library(tidyverse)
library(plotly)
```

<style>
.navbar{
background-color: #1380A1;
border-color: #1380A1
}

.navbar-inverse .navbar-nav > li > a:hover,
.navbar-inverse .navbar-nav > li > a:focus{
  background-color: #FAAB18;
  color: white;
}

.navbar-inverse .navbar-nav > .active > a,
.navbar-inverse .navbar-nav > .active > a:hover,
.navbar-inverse .navbar-nav > .active > a:focus{
  color: white;
  background-color: #FAAB18
}

.navbar-inverse .navbar-toggle: hover,
.navbar-inverse .navbar-toggle: focus{
  background-color: #FAAB18
}
.navbar-inverse .navbar-collapse,
.navbar-inverse .navbar-form{
  border-color: #FAAB18;
}
</style>
```{r include=FALSE}
init_data <- read_csv("Teenage Pregnancies Data 2016 to 2020_WardLevel_Quarterly.csv")

# How many adolescent pregnancies per district from 2016 - 2020
annual_preg <- init_data %>% 
  select(periodid, orgunitlevel2, 
         `adolescents (10-14 years) with  pregnancy`,
         `adolescents (15-19 years) with  pregnancy`) %>% 
  drop_na() %>% 
  glimpse()

which(is.na(annual_preg$`adolescents (15-19 years) with  pregnancy`))

annual_preg_1 <- annual_preg %>% 
  mutate(year = substring(periodid,1,4)) %>% 
  mutate(q = substring(periodid,6,6)) %>% 
  glimpse()

annual_preg_2 <- annual_preg_1 %>% 
  mutate(year_q = paste0(year,".", q )) %>% 
  glimpse()

annual_preg_2$year_q <- as.numeric(annual_preg_2$year_q)

str(annual_preg_2)
range(annual_preg_2$year_q)

annual_preg_2 <- annual_preg_2 %>% 
  arrange(year_q) %>% 
  glimpse()

annual_preg_2 %>% distinct(orgunitlevel2) %>% 
  count()

str(annual_preg_2)

annual_preg_3 <- annual_preg_2 %>% 
  select(year_q, orgunitlevel2, `adolescents (10-14 years) with  pregnancy`,`adolescents (15-19 years) with  pregnancy`) %>%
  group_by(orgunitlevel2, year_q) %>% 
  arrange(year_q) %>%
  mutate(total = `adolescents (10-14 years) with  pregnancy`+`adolescents (15-19 years) with  pregnancy`) 

annual_preg_4 <- annual_preg_3 %>% 
  select(year_q, orgunitlevel2, total) %>% 
  rename(Year = year_q, County = orgunitlevel2)

annual_preg_5 <-
  summarise(annual_preg_4,
            total = sum(total))

summary(annual_preg_3)

summary(annual_preg_5)

library(viridis)
library(hrbrthemes)

pt1 <- annual_preg_5 %>%
  ggplot( aes(x=Year, y=total, group=County, color = County)) +
  geom_line() +
  #scale_color_manual(values = c("#69b3a2", "lightgrey")) +
  #scale_size_manual(values=c(1.5,0.2)) +
  theme(legend.position="none") +
  ggtitle("Adolescent Pregnancies in Kenya by County") +
  theme_ipsum() +
  xlab("Year") +
  ylab("Total pregnancies") +
  #geom_label( x=1990, y=55000, label="Amanda reached 3550\nbabies in 1970", size=4, color="#69b3a2") +
  theme(
    legend.position="none",
    plot.title = element_text(size=14)
  )

#facet_wrap(~district)
library(Cairo)

ggplotly(pt1)

init_data %>% distinct(orgunitlevel4) %>% 
  count()

library("ggalt")
library("tidyr")
library(bbplot)

#Prepare data
dumbbell_df <- annual_preg_5 %>%
  #  filter(Year == 2016.1 | 
  #           Year == 2020.2 |
  #           Year == 2016.1) %>%
  select(County, Year, total) %>%
  spread(Year, total) %>%
  replace_na(list(`2016.1` = 0, `2016.2` = 0,`2016.3` = 0,`2016.4` = 0,
                  `2017.1` = 0, `2017.2` = 0,`2017.3` = 0,`2017.4` = 0,
                  `2018.1` = 0, `2018.2` = 0,`2018.3` = 0,`2018.4` = 0,
                  `2019.1` = 0, `2019.2` = 0,`2019.3` = 0,`2019.4` = 0,
                  `2020.1` = 0, `2020.2` = 0)) %>% 
  mutate(gap = `2020.2` - `2016.1`) %>%
  arrange(desc(gap)) %>%
  head(47)

unique(dumbbell_df$County)

# Modification to axis test size of the bbstyle() function to use for overview page
use_this_style <- function(){
  font <- "Helvetica"
  ggplot2::theme(plot.title = ggplot2::element_text(family = font, 
                                                    size = 12, face = "bold", color = "#222222"),
                 plot.caption = ggplot2::element_text(family = font, 
                                                    size = 8, color = "#222222"),
                 plot.subtitle = ggplot2::element_text(family = font, 
                                                      size = 22, margin = ggplot2::margin(9, 0, 9,0)),
                 legend.position = "top", legend.text.align = 0,
                 legend.title = ggplot2::element_blank(), 
                 legend.text = ggplot2::element_text(family = font, size = 10, 
                                                     color = "#222222"), axis.title = ggplot2::element_blank(), 
                 axis.text = ggplot2::element_text(family = font, size = 10, 
                                                   color = "#222222"), axis.text.x = ggplot2::element_text(margin = ggplot2::margin(5, b = 10)), axis.ticks = ggplot2::element_line(color = "#cbcbcb"), 
                 axis.line = ggplot2::element_line(color = "#cbcbcb"), panel.grid.minor = ggplot2::element_blank(), 
                 panel.grid.major.y = ggplot2::element_line(color = "#cbcbcb"), 
                 panel.grid.major.x = ggplot2::element_blank(), panel.background = ggplot2::element_blank(),strip.background = ggplot2::element_rect(fill = "white"), 
                 strip.text = ggplot2::element_text(size = 22, hjust = 0))
}

# Modification to axis test size of the bbstyle() function to use for detail page
use_this_style2 <- function(){
  font <- "Helvetica"
  ggplot2::theme(plot.title = ggplot2::element_text(family = font, 
                                                    size = 12, face = "bold", color = "#222222"),
                 plot.caption = ggplot2::element_text(family = font, 
                                                    size = 10, color = "#222222"),
                 plot.subtitle = ggplot2::element_text(family = font, 
                                                      size = 22, margin = ggplot2::margin(9, 0, 9,0)),
                 legend.position = "top", legend.text.align = 0,
                 legend.title = ggplot2::element_blank(), 
                 legend.text = ggplot2::element_text(family = font, size = 12, 
                                                     color = "#222222"), axis.title = ggplot2::element_blank(), 
                 axis.text = ggplot2::element_text(family = font, size = 12, 
                                                   color = "#222222"), 
                 axis.text.x = ggplot2::element_text(family = font, size = 12,color = "#222222"), 
                 axis.ticks = ggplot2::element_line(color = "#cbcbcb"), 
                 axis.line = ggplot2::element_line(color = "#cbcbcb"), panel.grid.minor = ggplot2::element_blank(), 
                 panel.grid.major.y = ggplot2::element_line(color = "#cbcbcb"), 
                 panel.grid.major.x = ggplot2::element_line(color = "#cbcbcb"), 
                 panel.background = ggplot2::element_blank(),strip.background = ggplot2::element_rect(fill = "white"), 
                 strip.text = ggplot2::element_text(size = 22, hjust = 0))
}

#Make plot for overview page
db_plot <- ggplot(dumbbell_df, aes(x = `2016.1`, xend = `2020.2`, y = reorder(County, gap), group = County)) + 
  geom_dumbbell(colour = "#dddddd",
                size = 1.8,
                colour_x = "#FAAB18",
                colour_xend = "#1380A1"
                ) +
  use_this_style() + 
  labs(title="Rise in reported pregnancies per county, 2016-2020",
       caption = "Data source: Women in GIS, Kenya and DHIS2 July 2020")

glimpse(annual_preg_3)
library(GGally)
library(ggplot2)

# ages 10 - 14
ages_10_14 <- annual_preg_3 %>% 
  select(year_q, orgunitlevel2, `adolescents (10-14 years) with  pregnancy`)

ages_10_14 <- ages_10_14 %>%
  gather("age_group", "count", 3)

# ages 15 - 19
ages_15_19 <- annual_preg_3 %>% 
  select(year_q, orgunitlevel2, `adolescents (15-19 years) with  pregnancy`)

ages_15_19 <- ages_15_19 %>%
  gather("age_group", "count", 3)

# combine ages 10-14 with ages 15-19
pregnancy_groups <- bind_rows(ages_10_14,ages_15_19) %>% 
  dplyr::group_by(orgunitlevel2) %>% 
  dplyr::arrange(year_q) %>% 
  dplyr::rename(Year = year_q, County = orgunitlevel2, Total = count, Group = age_group)

#Make plot
grouped_bars <- ggplot(pregnancy_groups, 
                       aes(x = Year, 
                           y = Total, 
                           fill = as.factor(Group))) +
  geom_bar(stat="identity", position="dodge") +
  geom_hline(yintercept = 0, size = 1, colour="#333333") +
  geom_hline(yintercept = mean(pregnancy_groups$Total), size = 1,
             lty = 2, colour = "red") +
  bbc_style() +
  scale_fill_manual(values = c("#1380A1", "#FAAB18")) +
  labs(title="We're living longer",
       subtitle = "Biggest life expectancy rise, 1967-2007") +
  facet_wrap(~ County)

min_mean_pregnancies <- pregnancy_groups[which.min(pregnancy_groups$Total),]
max_mean_pregnancies <- pregnancy_groups[which.min(pregnancy_groups$Total),]
glimpse(init_data)

detail_data <- init_data %>% 
  select(periodid, orgunitlevel2, orgunitlevel3, orgunitlevel4,
         `adolescents (10-14 years) with  pregnancy`,
         `adolescents (15-19 years) with  pregnancy`) %>% 
  mutate(sub_county = str_sub(orgunitlevel3, end = -11)) %>% 
  mutate(ward = str_sub(orgunitlevel4, end = -5))

# replace NAs with 0
detail_data <- detail_data %>% 
  rename(`Age(10 - 14)` = `adolescents (10-14 years) with  pregnancy`,
         `Age(15 - 19)` = `adolescents (15-19 years) with  pregnancy`,
         county = orgunitlevel2) %>% 
  replace_na(list(`Age(10 - 14)` = 0, `Age(15 - 19)` = 0))

detail_data_1 <- detail_data %>% 
  mutate(year = substring(periodid,1,4)) %>% 
  mutate(q = substring(periodid,6,6)) %>% 
  glimpse()

detail_data_2 <- detail_data_1 %>% 
  mutate(Year_quarter = paste0(year,".", q ))

detail_data_2 <- detail_data_2 %>% 
  select(-c(periodid, year, q, orgunitlevel3, orgunitlevel4))

detail_data_2$Year_quarter <- as.numeric(detail_data_2$Year_quarter)

str(detail_data_2)

detail_data_2 <- detail_data_2 %>% 
  mutate(total = `Age(10 - 14)` + `Age(15 - 19)`)

max(detail_data_2$total)
range(detail_data_2$total)

#Make plot
grouped_bars <- ggplot(pregnancy_groups, 
                       aes(x = Year, 
                           y = Total, 
                           fill = as.factor(Group))) +
  geom_bar(stat="identity", position="dodge") +
  geom_hline(yintercept = 0, size = 1, colour="#333333") +
  geom_hline(yintercept = mean(pregnancy_groups$Total), size = 1,
             lty = 2, colour = "red") +
  use_this_style() +
  scale_fill_manual(values = c("#1380A1", "#FAAB18")) +
  labs(title="We're living longer",
       subtitle = "Biggest life expectancy rise, 1967-2007") +
  facet_wrap(~ County)

# detail tabset 2: uptake of family planning
family_plan <- init_data %>% 
  select(periodid, orgunitlevel2, orgunitlevel3, orgunitlevel4, `adolescents (10-14 years) with  pregnancy`, `adolescent family planning uptake 10-14 yrs`, `adolescents (15-19 years) with  pregnancy`, `adolescent family planning uptake 15-19 yrs`)

family_plan_1 <- family_plan %>% 
  rename(`Age(10 - 14)` = `adolescents (10-14 years) with  pregnancy`,
         `Age(15 - 19)` = `adolescents (15-19 years) with  pregnancy`,
         county = orgunitlevel2, sub_county = orgunitlevel3, ward = orgunitlevel4,
         `Family planning uptake: Age(10 - 14)` = `adolescent family planning uptake 10-14 yrs`,
         `Family planning uptake: Age(15 - 19)` = `adolescent family planning uptake 15-19 yrs`) %>% 
  replace_na(list(`Age(10 - 14)` = 0, `Age(15 - 19)` = 0, `Family planning uptake: Age(10 - 14)` = 0,
                  `Family planning uptake: Age(15 - 19)` = 0)) %>% 
  mutate(sub_county = str_sub(sub_county, end = -11)) %>% 
  mutate(ward = str_sub(ward, end = -5)) %>% 
  mutate(year = substring(periodid,1,4)) %>% 
  mutate(q = substring(periodid,6,6)) %>% 
  mutate(Year_quarter = paste0(year,".", q ))

family_plan_2 <- family_plan_1 %>% 
  select(-c(periodid,year, q))

family_plan_3 <- family_plan_2 %>% 
  select(Year_quarter, county, sub_county, ward,`Age(10 - 14)`, `Age(15 - 19)`,
         `Family planning uptake: Age(10 - 14)`,`Family planning uptake: Age(15 - 19)`)

family_plan_4 <- gather(family_plan_3, "Age Group", "Pregnancies", 5:6)
family_plan_5 <- gather(family_plan_4, "Age Group Family Planning", "Uptake of Family Planning", 5:6)
family_plan_5$Year_quarter <- as.numeric(family_plan_5$Year_quarter)

family_plan_6 <- family_plan_5 %>% 
  select(-`Age Group Family Planning`)

family_plan_7 <- gather(family_plan_6, "Category", "Total", 6:7)
family_plan_7$Category <- as_factor(family_plan_7$Category)

```

Overview
=======================================================================

Row
-----------------------------------------------------------------------


```{r fig.width=15, fig.height = 10}
(db_plot)
```

Detail
=======================================================================

Sidebar {.sidebar}
-----------------------------------------------------------------------
```{r fig.width=12}
# shiny inputs defined here
counties <- c("Narok County", "Homa Bay County", "Bungoma County", "Nairobi County",
              "Trans Nzoia County", "Kajiado County", "West Pokot County", "Meru County",
              "Siaya County", "Kakamega County", "Kwale County", "Turkana County",
              "Kisii County", "Nakuru County","Nyamira County","Bomet County","Kericho County",
              "Kisumu County","Vihiga County", "Kilifi County", "Samburu County", "Makueni County",
              "Baringo County","Kitui County","Tana River County","Laikipia County",
              "Elgeyo Marakwet County","Marsabit County","Isiolo County","Nandi County",
              "Uasin Gishu County","Muranga County","Mandera County","Kiambu County",
              "Tharaka Nithi County","Nyandarua County","Machakos County","Taita Taveta County",
              "Kirinyaga County","Embu County","Mombasa County","Busia County","Lamu County",
              "Garissa County","Migori County","Nyeri County","Wajir County")

fillCol(height = 600, flex = c(NA, 1), 
    inputPanel(
      radioButtons("county", "Select county:", choices = counties))
    )  
```

Row
-----------------------------------------------------------------------

### Year & quarter with highest reported cases 
```{r }
#detail_data_2$sub_county <-unlist(strsplit(detail_data_2$sub_county, split=' ', fixed=TRUE))[1]
#detail_data_2$ward <-unlist(strsplit(detail_data_2$ward, split=' ', fixed=TRUE))[1]

selectedBox1 <- reactive({detail_data_2 %>%
                        dplyr::filter(county == input$county)})

renderValueBox({
  display_year <- selectedBox1()[which.max(selectedBox1()$total),]
  
  valueBox(value = display_year$Year_quarter,
           icon = "fa-chart-line",
           color = "#1380A1"
           )
})
```

### Sub-County with highest reported cases
```{r}
selectedBox2 <- reactive({detail_data_2 %>%
                        dplyr::filter(county == input$county)})

renderValueBox({
  display_sub_county <- selectedBox2()[which.max(selectedBox1()$total),]
  
  valueBox(value = display_sub_county$sub_county,
           icon = "fa-venus",
           color = "#FAAB18"
           )
})
```

### Ward with highest reported cases
```{r}
selectedBox3 <- reactive({detail_data_2 %>%
                        dplyr::filter(county == input$county)})

renderValueBox({
  display_sub_county <- selectedBox3()[which.max(selectedBox1()$total),]
  
  valueBox(value = display_sub_county$ward,
           icon = "fa-female",
           color = "#1380A1"
           )
})
```

Row
-----------------------------------------------------------------------
```{r}
detail_data_3 <- detail_data_2 %>% 
  select(county, sub_county, ward, `Age(10 - 14)`, `Age(15 - 19)`, Year_quarter)

detail_data_4 <- gather(detail_data_3, "Age Group", "Total",4:5)
  

selectedCounty <- reactive({detail_data_4 %>%
                        dplyr::filter(county == input$county)})

#Make plot

renderPlot({ggplot(selectedCounty(), 
         aes(x = selectedCounty()$Year_quarter, y = selectedCounty()$Total, 
                           fill = as.factor(selectedCounty()$`Age Group`))) +
  geom_bar(stat="identity", position = "dodge") +
  geom_hline(yintercept = 0, size = 1, colour="#333333") +
  #geom_hline(yintercept = mean(selectedCounty()$Total), size = 1,
         #    lty = 2, colour = "red") +
  use_this_style2() +
  scale_fill_manual(values = c("#1380A1", "#FAAB18")) +
  labs(
       subtitle = "Adolescent pregnancies per yearly quarter",
       caption = "Data source: Women in GIS, Kenya and DHIS2 July 2020") +
  theme(legend.position = "bottom",
        axis.text.x = element_text(face="bold", color="#333333", 
                           size=14),
        axis.ticks.x = element_line(color = "#cbcbcb"))})
```

Family planning effect
=======================================================================

Sidebar {.sidebar}
-----------------------------------------------------------------------
```{r}
age <- c("Age(10 - 14)", "Age(15 - 19)")

      selectInput("countyFP", "Select county:", choices = counties)
      selectInput("age", "Select age group:", choices = age)
```

```{r}
selectedDATA <- reactive({family_plan_7 %>%
                        dplyr::filter(county == input$countyFP & age == input$age)
                        })
```
Row
-----------------------------------------------------------------------

### Year & quarter where family planning had the greatest effect
```{r }
selectedBox4 <- reactive({family_plan_6 %>%
                        dplyr::filter(county == input$countyFP & age == input$age)})

renderValueBox({
  display_year <- selectedBox4()[which.max(selectedBox4()$`Uptake of Family Planning` - selectedBox4()$`Pregnancies`),]
  
  valueBox(value = display_year$Year_quarter,
           icon = "fa-chart-line",
           color = "#1380A1"
           )
})
```

### Sub-County where family planning had the greatest effect
```{r}
selectedBox5 <- reactive({family_plan_6 %>%
                        dplyr::filter(county == input$countyFP & age == input$age)})

renderValueBox({
  display_sub_county <- selectedBox5()[which.max(selectedBox5()$`Uptake of Family Planning` - selectedBox4()$`Pregnancies`),]
  
  valueBox(value = display_sub_county$sub_county,
           icon = "fa-venus",
           color = "#FAAB18"
           )
})
```

### Ward where family planning had the greatest effect
```{r}
selectedBox6 <- reactive({family_plan_6 %>%
                        dplyr::filter(county == input$countyFP & age == input$age)})

renderValueBox({
  display_sub_county <- selectedBox6()[which.max(selectedBox6()$`Uptake of Family Planning` - selectedBox6()$`Pregnancies`),]
  
  valueBox(value = display_sub_county$ward,
           icon = "fa-female",
           color = "#1380A1"
           )
})
```

Row
-----------------------------------------------------------------------

```{r}
#Make plot
quarter_year <- c("2016.1","2016.2", "2016.3", "2016.4",
                  "2017.1","2017.2", "2017.3", "2017.4",
                  "2018.1","2018.2", "2018.3", "2018.4",
                  "2019.1","2019.2", "2019.3", "2019.4",
                  "2020.1","2020.2")

renderPlot({ggplot(selectedDATA(), 
         aes(x = selectedDATA()$Year_quarter, y = selectedDATA()$Total, 
                           fill = selectedDATA()$Category)) +
  geom_bar(stat="identity", position = "dodge") +
  geom_hline(yintercept = 0, size = 1, colour="#333333") +
  #geom_hline(yintercept = mean(selectedCounty()$Total), size = 1,
         #    lty = 2, colour = "red") +
  use_this_style2() +
  scale_fill_manual(values = c("#1380A1", "#FAAB18")) +
  labs(
       subtitle = "Effect of family planning on adolescent pregnancies per yearly quarter",
       caption = "Data source: Women in GIS, Kenya and DHIS2 July 2020") +
  theme(legend.position = "bottom",
        axis.text.x = element_text(face="bold", color="#333333", 
                           size=14),
        axis.ticks.x = element_line(color = "#cbcbcb"))})
```

